Competing Bandits: The Perils of Exploration under Competition

### 2018-09-02 paper/content/barriers.tex
basically rewrite barriers section, tweaks to inverted U and sim details

An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant. This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant. Which plays a bigger role in preventing the entrant from being able to establish market share? We run two additional experiments, modifying the previous temporary monopoly experiment so that in one experiment the reputation of the incumbent is artificially erased (Reputation Erased Experiment) and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).

@@

\begin{finding}
\textit{Both the data advantage and reputation advantage alone serve as strong barriers to entry. When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
\end{finding}

Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments. Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market. The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $\TS$. One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $\TS$ is sufficient so that it becomes relatively easier for $\TS$ to recover the reputation loss compared to the information loss, but since $\DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $\TS$ there is less of an informational advantage.

Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained. We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.

>>>>

An alternative interpretation of the results in the previous section is that the temporary monopoly provides a first mover advantage to the incumbent firm. The early entry allows the incumbent to acquire data on potentially suboptimal actions without incurring an immediate reputational consequence as well as the ability to eventually build up a high reputation before the entrant enters the market. Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant. This gives new intuition for the classic ``first mover advantage" \cite{kerin1992first} in the digital economy where data can serve as a barrier to entry in many markets.

While combined these two advantages serve as strong barriers to entry a natural question to ask is would having only the data advantage or only the reputation advantage still serve as such strong barriers to entry? Which of the two serves as a stronger barrier to entry and under what conditions? In order to answer these questions we run two additional experiments, artificially modifying the previous temporary monopoly experiment. In one experiment the reputation of the incumbent is artificially erased upon the entry of the entrant (Reputation Erased Experiment) and in the other the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).

@@

\begin{finding}
\textit{Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market. \\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two.}
\end{finding}

Tables \ref{info_erase} and \ref{rep_erase} display the results of these experiments. The reason for the finding is that, under $\TS$, the early entry allows the incumbent to do exploration without immediate reputational consequence in the early rounds and so, even when reputation is erased, the incumbent has already acquired sufficient information that simply exploiting that information allows her to quickly recover her reputation. However, while the reputational advantage gives the incumbent some cushion for doing exploration the incumbent is still vulnerable to getting unlucky while doing exploration and suffering the reputational consequences. This only happens under $\TS$ and not under $\DEG$ or $\DG$ since $\TS$, by construction, is more intelligent about the data that it acquires compared to $\DEG$ and $\DG$.

Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained. We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.

### 2018-09-02 paper/content/inverted_u.tex
basically rewrite barriers section, tweaks to inverted U and sim details

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}
\section{Competition vs Better Algorithms}\label{section:5}
\swedit{how about just name the section ``inverted U''?}  \swedit{Now
  we demonstrate the inverted-U relationship between the incentivized
  algorithms and the competition levels. In particular, we investigate
  the incentivized algorithms under three different regimes: permanent
  monopoly, temporary monopoly, and permanent duopoly. First, the
  story for permanent monopoly is relatively straightforward.}

% couldn't get section references to work so hardcoding for now
We utilize the same instances and realizations as in the section 3  and simulate the model described in section 2. We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms. Unless otherwise noted, all the results are reported at $t = 2000$.
}

@@

\xhdr{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously. What algorithms are firms incentivized to deploy? We report the average market share taken over the $N$ simulations for a set of exogenous values of $T_0$. Additionally, we report the mean and median of an additional quantity which we define as the \textit{effective end of game}.

@@

\begin{finding}
\textit{
\begin{enumerate}
\item For low warm starts, under Heavy Tail and Uniform \\ $\DG > \TS$ in competition (see Table \ref{sim_ht}),
\item Under Needle In Haystack, \\ $\TS > \DG$ (see Table \ref{sim_nih}). 
\item For sufficiently large warm starts, $\TS$ is dominant across all the families of instances.
\item Even for a large warm start, $\DG > \DEG$ under Heavy Tail
\end{enumerate}
}
\end{finding}

@@

In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail family \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$. However, while $\DG > \TS$ for the entrant, $\DG$ does roughly the same as $\DEG$ on the Uniform family. The results are in the supplementary material.}. In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the family of instances we consider.

@@

Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was preferred to $\TS$? The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds. One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation. The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition. \OMIT{As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.} Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.

>>>>

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}
\section{Competition vs Better Algorithms}\label{section:5}
\swedit{how about just name the section ``inverted U''?}  \swedit{Now
  we demonstrate the inverted-U relationship between the incentivized
  algorithms and the competition levels. In particular, we investigate
  the incentivized algorithms under three different regimes: permanent
  monopoly, temporary monopoly, and permanent duopoly. \gaedit{We break indifference between exploration algorithms and non-exploration algorithms towards non-exploration algorithms.} First, the story for permanent monopoly is relatively straightforward.}

@@

\xhdr{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously. What algorithms are firms incentivized to deploy? We report the average market share taken over the $N$ simulations for a set of exogenous values of $T_0$. Additionally, we report the mean and median of the effective end of game.

@@

\begin{finding}
\textit{
\begin{enumerate}
\item For low warm starts, under Heavy Tail and Uniform \\ $\DG > \TS$ in competition (see Table \ref{sim_ht}),
\item Under Needle In Haystack, \\ $\TS > \DG$ (see Table \ref{sim_nih}). 
\item For sufficiently large warm starts, $\TS$ is dominant across all instances.
\item Even for a large warm start, $\DG > \DEG$ under Heavy Tail
\end{enumerate}
}
\end{finding}

@@

In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail family \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.}. In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the family of instances we consider.

@@

Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was preferred to $\TS$? The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds. One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation. The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.

\OMIT{As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.} Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
